---
title: "Hot Spots Code"
author: "Martin Mokros"
date: "12/07/2020"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

Firstly we will import all the raster layers

four periods: 1961-1990 (past - p)   
              2041-2060 (Future first period - f1)  
              2061-2080 (Future second period - f2)   
              2081-2100 (Future third period - f3). 

nine variables: Cont; DDa5; EQ; FFP; LLDP; MWMT; nDP; pr_ann; tasmax

one model: CNRM

Altogether 36 raster layers will be diveded to four groups based on the time period


```{r}
library(raster)
library(rasterVis)
library(mgcv)
```

```{r}
dataroot = "C:/DATA_CNRM"

p.list = list.files(dataroot, pattern="1961",full.names=T)
f1.list = list.files(dataroot, pattern="2041",full.names=T)
f2.list = list.files(dataroot, pattern="2061",full.names=T)
f3.list = list.files(dataroot, pattern="2081",full.names=T)

rp.list = lapply(p.list, raster)
rf1.list = lapply(f1.list, raster)
rf2.list = lapply(f2.list, raster)
rf3.list = lapply(f3.list, raster)


for (i in 1:9) {
  NAvalue(rp.list[[i]]) = -3.402823e+38
  NAvalue(rf1.list[[i]]) = -3.402823e+38
  NAvalue(rf2.list[[i]]) = -3.402823e+38
  NAvalue(rf3.list[[i]]) = -3.402823e+38
}


```

Example of Cont for four periods from CNRM model:

```{r, echo=FALSE, fig.height= 15, fig.width= 15}

s = stack(rp.list[[1]], rf1.list[[1]], rf2.list[[1]], rf3.list[[1]])
levelplot(s, layout=c(2,2), par.settings=BuRdTheme, main="Cont", names.attr=c("1961-1990", "2041-2060", "2060-2081", "2081-2100"))


```

Calculation of Standardize Euclidean Distance (SED)

```{r}
SED1_list = list()
SED2_list = list()
SED3_list = list()

for (i in 1:9){

  p = rp.list[[i]]
  f1 = rf1.list[[i]]
  f2 = rf2.list[[i]]
  f3 = rf3.list[[i]]
  
  minus1 = abs(f1 - p)
  minus2 = abs(f2 - p)
  minus3 = abs(f3 - p)
  
  q1 = quantile(minus1,  probs = c(0.95))
  q2 = quantile(minus2,  probs = c(0.95))
  q3 = quantile(minus3,  probs = c(0.95))

  SED1 = (minus1 / q1)^2
  SED2 = (minus2 / q2)^2
  SED3 = (minus3 / q3)^2

  SED1[SED1>1] = 1
  SED2[SED2>1] = 1
  SED3[SED3>1] = 1
  
  SED1_list[[i]] = SED1
  SED2_list[[i]] = SED2
  SED3_list[[i]] = SED3
}

SED1r = stack(SED1_list)
SED2r = stack(SED2_list)
SED3r = stack(SED3_list)

```

Example of SED for each time period

```{r, echo=FALSE, fig.height= 10, fig.width= 20}
s = stack(SED1r[[1]], SED2r[[1]], SED3r[[1]])
levelplot(s, layout=c(3,1), par.settings=BuRdTheme, main="Standardized Euclidean Distance", names.attr=c("2041-2060", "2060-2081", "2081-2100"))
```

Calculation of Aggregate Climate Change 

```{r}
CNRM_1 = sqrt(calc(subset(SED1r, 1:9), sum))
CNRM_2 = sqrt(calc(subset(SED2r, 1:9), sum))
CNRM_3 = sqrt(calc(subset(SED3r, 1:9), sum))

```

Visualization of Aggregate Climate Change for each time period and model

```{r, echo=FALSE, fig.height= 10, fig.width= 20}
s = stack(CNRM_1, CNRM_2, CNRM_3)
levelplot(s, layout=c(3,1), par.settings=BuRdTheme, main="Aggregate Climate Change",
          names.attr=c("CNRM 2041-2060", "CNRM 2061-2080", "CNRM 2081-2100"))
```


Calculation of relative Aggregate Climate Change (%) 

```{r}
CNRM_1_per = (CNRM_1/sqrt(9))*100
CNRM_2_per = (CNRM_2/sqrt(9))*100
CNRM_3_per = (CNRM_3/sqrt(9))*100
```

Visualization of relative Aggregate Climate Change (%) for each time period and model

```{r, echo=FALSE, fig.height= 10, fig.width= 20}
s = stack(CNRM_1_per, CNRM_2_per, CNRM_3_per)
levelplot(s, layout=c(3,1), par.settings=BuRdTheme, main="relative Aggregate Climate Change (%)",
          names.attr=c("CNRM 2041-2060", "CNRM 2061-2080", "CNRM 2081-2100"))

```

Applying generalized additive model (GAM) to identify regional HotSPots
```{r}
CNRM_1_gam = as.data.frame(CNRM_1_per, xy = TRUE)
CNRM_2_gam = as.data.frame(CNRM_2_per, xy = TRUE)
CNRM_3_gam = as.data.frame(CNRM_3_per, xy = TRUE)

la = CNRM_1_gam$y
lo = CNRM_1_gam$x
ACC = CNRM_1_gam$layer
M1 = gam(ACC~s(la,lo, bs="sos"))
sink("CNRM_1_gam.txt")
print(summary(M1))
sink()
CNRM_1_gam = na.omit(CNRM_1_gam)
CNRM_1_gam$residuals = residuals(M1)
CNRM_1_gam$fitted = fitted(M1)

la = CNRM_2_gam$y
lo = CNRM_2_gam$x
ACC = CNRM_2_gam$layer
M2 = gam(ACC~s(la,lo, bs="sos"))
sink("CNRM_2_gam.txt")
print(summary(M2))
sink()
CNRM_2_gam = na.omit(CNRM_2_gam)
CNRM_2_gam$residuals = residuals(M2)
CNRM_2_gam$fitted = fitted(M2)

la = CNRM_3_gam$y
lo = CNRM_3_gam$x
ACC = CNRM_3_gam$layer
M3 = gam(ACC~s(la,lo, bs="sos"))
sink("CNRM_3_gam.txt")
print(summary(M3))
sink()
CNRM_3_gam = na.omit(CNRM_3_gam)
CNRM_3_gam$residuals = residuals(M3)
CNRM_3_gam$fitted = fitted(M3)

CNRM_1_gam_brick = rasterFromXYZ(CNRM_1_gam)
CNRM_1_res = raster(CNRM_1_gam_brick, layer=2)
CNRM_1_fit = raster(CNRM_1_gam_brick, layer=3)

CNRM_2_gam_brick = rasterFromXYZ(CNRM_2_gam)
CNRM_2_res = raster(CNRM_2_gam_brick, layer=2)
CNRM_2_fit = raster(CNRM_2_gam_brick, layer=3)

CNRM_3_gam_brick = rasterFromXYZ(CNRM_3_gam)
CNRM_3_res = raster(CNRM_3_gam_brick, layer=2)
CNRM_3_fit = raster(CNRM_3_gam_brick, layer=3)

```

Visualization of calculated residuals by generalized additive model (GAM)
```{r, echo=FALSE, fig.height= 10, fig.width= 20}
s = stack(CNRM_1_res, CNRM_2_res, CNRM_3_res)
levelplot(s, layout=c(3,1), par.settings=BuRdTheme, main="Residuals",
          names.attr=c("CNRM 2041-2060", "CNRM 2061-2080", "CNRM 2081-2100"))

```

